Communicating human related spatial distributions: A comparison of spatial visualisations using visual inference.

### 2020-01-05 paper/paper.tex
paper changes

\hypersetup{
            pdftitle={Which is Better: a Choropleth Map or Hexagon Tile Map? A comparison using visual inference},
            pdfkeywords={statistics, visual inference, geospatial, population},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls

@@

\begin{document}
%
% paper title
% Titles are generally capitalized except for words such as a, an, and, as,
% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
% not capitalized unless they are the first or last word of the title.
% Linebreaks \\ can be used within to get better formatting as desired.
% Do not put math or special symbols in the title.
\title{Which is Better: a Choropleth Map or Hexagon Tile Map? A comparison
using visual inference}

@@

\hypertarget{australian-cancer-atlas}{%
\subsection{Australian Cancer Atlas}\label{australian-cancer-atlas}}

The Australian Cancer Atlas explores the burden of cancer on Australian
communities. There are many cancer types presented, and they can be
explored on an individual or aggregate level. The Australian communities
are examined at the Statistical Areas at Level 2 (SA2)(``Australian
Statistical Geography Standard (ASGS)''
\protect\hyperlink{ref-abs2016}{2018}) used by the Australian Bureau of
Statistics. Bayesian spatial smoothing has been applied to incorporate
the statistics of neighbouring areas, for both privacy and stability of
the estimates. The statistics that can be mapped are the diagnoses
(Standardised Incidence Rates) and excess deaths for each SA2,
communicated as the difference from the Australian average of the
statistics. The values of the statistic for each are communicated using
a diverging colour scheme.

@@

@@

Spatial visualisations - Choropleth However, the issue of using a
choropleth map base becomes obvious when considering the distributions.
Position is extremely important for analysis of a visualisation.

Map creators have the ability to present spatial statistics in
alternative displays that can highlight the population. This work aims
to show that a hexagon tile map display is a viable alternative to the
geographic map base for presenting population statistics. The same data
were shown on a choropleth map, and on a hexagon tile map. Comparing the
results of participants who see the choropleth to those who see a
hexagon tile map will show that population related distributions are
spotted more frequently in a hexagon tile map display.

@@

@@

The survey responses from participants were kept only if the participant
submitted answers for all 12 displays. This resulted in 95 participants.

@@

\begin{verbatim}
## # A tibble: 6 x 2
##   age     participants
##   <chr>          <int>
## 1 18 - 24           14
## 2 25 - 34           37
## 3 35 - 44           21
## 4 45 - 54           11
## 5 55+                6
## 6 NA                 3
\end{verbatim}

70 of the 95 participants were male, and 25 female and only two of the
participants had lived in Australia before.

@@

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/detection_compare-1.pdf}
\caption{Each point shows the probability of detection for the lineup
display, separated by the trend model hidden in the lineup. The points
for the same data set are linked to show the difference in the detection
rate when the same data was seen in each display. 11 of the 12 real
distribution plots were found more often in the hexagon display.}
\end{figure}

\begin{tabular}{l|l|r|r|r}
\hline
trend & type & replicate & mean & std.dev\\
\hline
NW-SE & Choropleth & 1 & 25.00 & 16.41\\
\hline
NW-SE & Choropleth & 2 & 21.12 & 15.28\\
\hline
NW-SE & Choropleth & 3 & 21.12 & 17.05\\
\hline
NW-SE & Choropleth & 4 & 21.20 & 14.52\\
\hline
NW-SE & Hexagon tiles & 1 & 19.53 & 14.56\\
\hline
NW-SE & Hexagon tiles & 2 & 21.95 & 16.42\\
\hline
NW-SE & Hexagon tiles & 3 & 22.02 & 16.12\\
\hline
NW-SE & Hexagon tiles & 4 & 18.17 & 14.51\\
\hline
three cities & Choropleth & 1 & 15.30 & 13.88\\
\hline
three cities & Choropleth & 2 & 25.22 & 14.36\\
\hline
three cities & Choropleth & 3 & 22.49 & 15.87\\
\hline
three cities & Choropleth & 4 & 18.68 & 13.74\\
\hline
three cities & Hexagon tiles & 1 & 20.37 & 17.02\\
\hline
three cities & Hexagon tiles & 2 & 19.95 & 16.59\\
\hline
three cities & Hexagon tiles & 3 & 20.02 & 17.19\\
\hline
three cities & Hexagon tiles & 4 & 18.02 & 16.93\\
\hline
all cities & Choropleth & 1 & 20.14 & 16.56\\
\hline
all cities & Choropleth & 2 & 22.75 & 14.72\\
\hline
all cities & Choropleth & 3 & 19.88 & 15.03\\
\hline
all cities & Choropleth & 4 & 19.33 & 16.15\\
\hline
all cities & Hexagon tiles & 1 & 19.83 & 15.87\\
\hline
all cities & Hexagon tiles & 2 & 16.96 & 14.65\\
\hline
all cities & Hexagon tiles & 3 & 19.21 & 15.27\\
\hline
all cities & Hexagon tiles & 4 & 20.03 & 15.11\\
\hline
\end{tabular}

@@

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/time-1.pdf}
\caption{The time taken to evaluate each display is broken into five
second windows. The height of the histogram bars show how many
evaulations were submitted within each time window. The distributions
for the choropleth and hexagon tile maps are very similar. Both have a
large peak at 0-5 seconds, and then a secondary peak at 10-20 seconds.
No response took over one minute.}
\end{figure}

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/time_compare-1.pdf}
\caption{Each point shows the average time taken for participants to
evalute the lineup display, separated by the trend model hidden in the
lineup. The points for the same data set are linked to show the
difference in the avergae time taken when the same data was seen in each
display. There is a lot of variation in the time taken. The shortest
average time (15 seconds), and the longest average time (23 seconds)
both occured when evaluating three cities}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/unnamed-chunk-1-1.pdf}
\caption{The amount of times each level of certainty was chosen by
participants when viewing hexagon tile map or choropleth displays.
Participants were more likely to choose a high certainty when
considering a Choropleth map. The default certainty of 3 was chosen most
for the Hexagon tile map displays.}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/unnamed-chunk-2-1.pdf}
\caption{The amount of times each level of certainty was chosen by
participants. The columns shown whether a viewer correctly selected the
real trend data plot when viewing hexagon tile map or choropleth
displays.}
\end{figure}

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/unnamed-chunk-3-1.pdf}
\caption{The amount of times each level of certainty was chosen by
participants. The columns shown whether a viewer correctly selected the
real trend data plot when viewing hexagon tile map or choropleth
displays. The rows show the type of the trend model added in the real
data plot. The default certainty level of 3 was chosen most frequently
when incorrect. Then shown a NW-SE or all cities trend participants felt
more certain of their correct choice.}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/unnamed-chunk-4-1.pdf}
\caption{The most common reason for choice of plot when looking at each
trend model shown in Choropleth and Hexagon Tile maps. Clusters were the
most common reason when viewing a Hexagon Tile map, trend was the most
common choice for choropleth displays except for the all cities
display.}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/unnamed-chunk-5-1.pdf}
\caption{The probablity of detection acheived by the contributors in
each group is shown by the points. Group B has a larger range and a
smaller inter-quartile range. Group A and both had 3 people who did not
find any of the data maps in the displays.}
\end{figure}

@@

\includegraphics{paper_files/figure-latex/unnamed-chunk-6-1.pdf}

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/choices-1.pdf}
\caption{Each facet is associated with one lineup, the height of the
bars count the choices made by the participants considering each lineup.
The bars coloured with black outlines show the map which contained a
trend model, these are the correct choices. The numbers differentiate
the replicates of each trend model and type of map display. Participants
were able to select 0 to indicate they did not want to choose a map.}
\end{figure}

@@

\begin{verbatim}
## # A tibble: 6 x 5
##   term                                estimate std.error statistic p.value
##   <chr>                                  <dbl>     <dbl>     <dbl>   <dbl>
## 1 (Intercept)                           0.0217     0.147     0.147   0.883
## 2 typeHexagon tiles                     0.420      0.211     1.99    0.047
## 3 trendthree cities                    -3.25       0.412    -7.88    0    
## 4 trendall cities                      -1.24       0.229    -5.41    0    
## 5 typeHexagon tiles:trendthree cities   2.37       0.465     5.10    0    
## 6 typeHexagon tiles:trendall cities     1.08       0.312     3.47    0.001
\end{verbatim}

\begin{verbatim}
## Analysis of Deviance Table
## 
## Model: binomial, link: logit
## 
## Response: detect_f
## 
## Terms added sequentially (first to last)
## 
## 
##            Df Deviance Resid. Df Resid. Dev  Pr(>Chi)    
## NULL                        1103     1477.0              
## type        1   83.526      1102     1393.5 < 2.2e-16 ***
## trend       2  101.699      1100     1291.8 < 2.2e-16 ***
## type:trend  2   35.517      1098     1256.2 1.939e-08 ***
## ---
## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
\end{verbatim}

\begin{verbatim}
##            
##             Incorrect Correct
##   Incorrect       431     121
##   Correct         242     310
\end{verbatim}

@@

\begin{verbatim}
## # A tibble: 8 x 5
##   term                                estimate std.error statistic group      
##   <chr>                                  <dbl>     <dbl>     <dbl> <chr>      
## 1 (Intercept)                            0.505    0.0340     14.9  fixed      
## 2 typeHexagon tiles                      0.103    0.0448      2.31 fixed      
## 3 trendthree cities                     -0.467    0.0448    -10.4  fixed      
## 4 trendall cities                       -0.277    0.0448     -6.19 fixed      
## 5 typeHexagon tiles:trendthree cities    0.250    0.0633      3.95 fixed      
## 6 typeHexagon tiles:trendall cities      0.239    0.0633      3.77 fixed      
## 7 sd_(Intercept).contributor             0.118   NA          NA    contributor
## 8 sd_Observation.Residual                0.430   NA          NA    Residual
\end{verbatim}

\begin{verbatim}
##            
##             Incorrect Correct
##   Incorrect        49       1
##   Correct         624     430
\end{verbatim}

@@

@@

\begin{verbatim}
## Linear mixed model fit by REML ['lmerMod']
## Formula: detect ~ type * trend + (1 | contributor)
##    Data: d
## REML criterion at convergence: 1350.719
## Random effects:
##  Groups      Name        Std.Dev.
##  contributor (Intercept) 0.1179  
##  Residual                0.4296  
## Number of obs: 1104, groups:  contributor, 92
## Fixed Effects:
##                         (Intercept)                    typeHexagon tiles  
##                              0.5054                               0.1033  
##                   trendthree cities                      trendall cities  
##                             -0.4674                              -0.2772  
## typeHexagon tiles:trendthree cities    typeHexagon tiles:trendall cities  
##                              0.2500                               0.2391
\end{verbatim}

@@

@@

>>>>

\hypersetup{
            pdftitle={Communicating human related spatial distributions: A comparison of spatial visualisations using visual inference.},
            pdfkeywords={statistics, visual inference, geospatial, population},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls

@@

\begin{document}
%
% paper title
% Titles are generally capitalized except for words such as a, an, and, as,
% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
% not capitalized unless they are the first or last word of the title.
% Linebreaks \\ can be used within to get better formatting as desired.
% Do not put math or special symbols in the title.
\title{Communicating human related spatial distributions: A comparison of
spatial visualisations using visual inference.}

@@

\hypertarget{the-australian-cancer-atlas}{%
\subsection{The Australian Cancer
Atlas}\label{the-australian-cancer-atlas}}

The Australian Cancer Atlas is an online interactive webtool created to
explore the burden of cancer on Australian communities. There are many
cancer types presented, and they can be explored on an individual or
aggregate level. The Australian communities are examined at the
Statistical Areas at Level 2 (SA2) (``Australian Statistical Geography
Standard (ASGS)'' \protect\hyperlink{ref-abs2016}{2018}) used by the
Australian Bureau of Statistics. Bayesian spatial smoothing has been
applied to incorporate the value of the statistics of the neighbouring
areas. Spatial smoothing allows for the protection privacy and gives
stability to the estimates. The statistics that can be mapped are the
diagnoses (Standardised Incidence Rates) and excess deaths for each SA2,
communicated as the difference from the Australian average of the
statistics. The values of the statistic for each geographic area is
communicated through the colour used to fill the areas, the colours are
chosen for the diverging colour scheme.

@@

Spatial visualisations communicate the distribution of statistics over
geographic space. The most common display for spatial data is the
choropleth map. However, the issue of using a choropleth map base
becomes obvious when considering the distributions. Map creators have
the ability to present spatial statistics in alternative displays that
can highlight the population. This work aims to show that a hexagon tile
map display is a viable alternative to the geographic map base for
presenting population statistics. The same data were shown on a
choropleth map, and on a hexagon tile map. Comparing the results of
participants who see the choropleth to those who see a hexagon tile map
will show that population related distributions are spotted more
frequently in a hexagon tile map display.

@@

@@

The mean of the detection rate for choropleth map, denoted as \(\mu_C\),
and the hexagon tile maps, \(\mu_H\) will be contrasted. This leads to
the following one sided hypothesis for this study:

\(H_0\) : \(\mu_H\) = \(\mu_C\) \(H_a\) : \(\mu_H\) \textgreater{}
\(\mu_C\)

The detection rate \(\hat\pi\) is calculated as the amount of people
that made the choice of plot that contained the real data, out of the
participants who saw data plot in the lineup of the null data plots.

the difference in the detection rates for the two displays will be
compared using the following:

\[ \hat\pi_C - \hat\pi_H \pm t_{1-\alpha,n-1}\sqrt{\hat\pi_C(1-\hat\pi_C)/n_C + \hat\pi_H(1-\hat\pi_H)/n_H} \]

@@

The survey responses from participants were kept only if the participant
submitted answers for all 12 displays. This resulted in 92 participants.

@@

67 of the 92 participants were male, and 25 female and only two of the
participants had lived in Australia before.

@@

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/detection_compare-1.pdf}
\caption{Contrasting the detection rate achieved by participants when
viewing the four replicates of the three trend models. Each point shows
the probability of detection for the lineup display, the facets separate
the trend models hidden in the lineup. The points for the same data set
shown in a choroleth or haxgon tile map display are linked to show the
difference in the detection rate.}
\end{figure}

\begin{tabular}{l|l|r|r|r}
\hline
trend & type & replicate & mean & std.dev\\
\hline
NW-SE & Choro. & 1 & 25.00 & 16.41\\
\hline
NW-SE & Choro. & 2 & 21.12 & 15.28\\
\hline
NW-SE & Choro. & 3 & 21.12 & 17.05\\
\hline
NW-SE & Choro. & 4 & 21.20 & 14.52\\
\hline
NW-SE & Hex. & 1 & 19.53 & 14.56\\
\hline
NW-SE & Hex. & 2 & 21.95 & 16.42\\
\hline
NW-SE & Hex. & 3 & 22.02 & 16.12\\
\hline
NW-SE & Hex. & 4 & 18.17 & 14.51\\
\hline
three cities & Choro. & 1 & 15.30 & 13.88\\
\hline
three cities & Choro. & 2 & 25.22 & 14.36\\
\hline
three cities & Choro. & 3 & 22.49 & 15.87\\
\hline
three cities & Choro. & 4 & 18.68 & 13.74\\
\hline
three cities & Hex. & 1 & 20.37 & 17.02\\
\hline
three cities & Hex. & 2 & 19.95 & 16.59\\
\hline
three cities & Hex. & 3 & 20.02 & 17.19\\
\hline
three cities & Hex. & 4 & 18.02 & 16.93\\
\hline
all cities & Choro. & 1 & 20.14 & 16.56\\
\hline
all cities & Choro. & 2 & 22.75 & 14.72\\
\hline
all cities & Choro. & 3 & 19.88 & 15.03\\
\hline
all cities & Choro. & 4 & 19.33 & 16.15\\
\hline
all cities & Hex. & 1 & 19.83 & 15.87\\
\hline
all cities & Hex. & 2 & 16.96 & 14.65\\
\hline
all cities & Hex. & 3 & 19.21 & 15.27\\
\hline
all cities & Hex. & 4 & 20.03 & 15.11\\
\hline
\end{tabular}

@@

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/hist_height-1.pdf}
\caption{The time taken to evaluate each display is broken into five
second windows. The height of the histogram bars show how many
evaulations were submitted within each time window. The distributions
for the choropleth and hexagon tile maps are very similar. Both have a
large peak at 0-5 seconds, and then a secondary peak at 10-20 seconds.
No response took over one minute.}
\end{figure}

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/hist_fill-1.pdf}
\caption{The time taken to evaluate each display is broken into five
second windows. The height of the histogram bars show the proportion of
evaulations that were submitted within each time window.}
\end{figure}

\includegraphics{paper_files/figure-latex/unnamed-chunk-1-1.pdf}

@@

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/unnamed-chunk-2-1.pdf}
\caption{The amount of times each level of certainty was chosen by
participants when viewing hexagon tile map or choropleth displays.
Participants were more likely to choose a high certainty when
considering a Choropleth map. The default certainty of 3 was chosen most
for the Hexagon tile map displays.}
\end{figure}

@@

@@

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/unnamed-chunk-3-1.pdf}
\caption{The most common reason for choice of plot when looking at each
trend model shown in Choropleth and Hexagon Tile maps. Clusters were the
most common reason when viewing a Hexagon Tile map, trend was the most
common choice for choropleth displays except for the all cities
display.}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/unnamed-chunk-4-1.pdf}
\caption{The probablity of detection acheived by the contributors in
each group is shown by the points. Group B has a larger range and a
smaller inter-quartile range. Group A and both had 3 people who did not
find any of the data maps in the displays.}
\end{figure}

@@

\includegraphics{paper_files/figure-latex/unnamed-chunk-5-1.pdf}

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/choices-1.pdf}
\caption{Each facet is associated with one lineup, the height of the
points show the proportion of the participants that made each choice
when considering each lineup. The points coloureddark blue show the map
which contained a trend model, these are the correct choices. The
numbers differentiate the replicates of each trend model and type of map
display. Participants were able to select 0 to indicate they did not
want to choose a map.}
\end{figure}

@@

\begin{tabular}{l|r|r|r|r|l}
\hline
\_ & terms & Estimate & Error & p.value & sig\\
\hline
Design & Choro & 0.0217400 & 0.1474507 & 0.8828 & \\
\hline
 & Hex & 0.4200928 & 0.2110904 & 0.0466 & *\\
\hline
Trend & Three cities & -3.2519793 & 0.4124809 & 0.0000 & ***\\
\hline
 & All cities & -1.2398974 & 0.2293327 & 0.0000 & ***\\
\hline
Interactions & Hex : Three cities & 2.3683138 & 0.4645163 & 0.0000 & ***\\
\hline
 & Hex : All cities & 1.0825772 & 0.3123987 & 0.0005 & ***\\
\hline
\end{tabular}

\begin{verbatim}
##      
##        No Yes
##   No  431 121
##   Yes 242 310
\end{verbatim}

This gives the model:

\[\widehat{y_{ij}} = 0.0217 + 0.42Hex -3.25_{three} -1.24_{all} + 2.37{Hex.three} + 1.08{Hex.all} \]

\includegraphics{paper_files/figure-latex/unnamed-chunk-8-1.pdf}

@@

\begin{verbatim}
## # A tibble: 8 x 5
##   term                       estimate std.error statistic group      
##   <chr>                         <dbl>     <dbl>     <dbl> <chr>      
## 1 (Intercept)                   0.505    0.0340     14.9  fixed      
## 2 typeHex.                      0.103    0.0448      2.31 fixed      
## 3 trendthree cities            -0.467    0.0448    -10.4  fixed      
## 4 trendall cities              -0.277    0.0448     -6.19 fixed      
## 5 typeHex.:trendthree cities    0.250    0.0633      3.95 fixed      
## 6 typeHex.:trendall cities      0.239    0.0633      3.77 fixed      
## 7 sd_(Intercept).contributor    0.118   NA          NA    contributor
## 8 sd_Observation.Residual       0.430   NA          NA    Residual
\end{verbatim}

\begin{verbatim}
##      
##        No Yes
##   No   49   1
##   Yes 624 430
\end{verbatim}

@@

\begin{verbatim}
## # A tibble: 1 x 6
##   sigma logLik   AIC   BIC deviance df.residual
##   <dbl>  <dbl> <dbl> <dbl>    <dbl>       <int>
## 1 0.445  -685. 1380. 1405.    1365.        1099
\end{verbatim}

\begin{verbatim}
## # A tibble: 5 x 5
##   term                           estimate std.error statistic group   
##   <chr>                             <dbl>     <dbl>     <dbl> <chr>   
## 1 (Intercept)                       0.669    0.0501      13.4 fixed   
## 2 sd_(Intercept).trend              0.456   NA           NA   trend   
## 3 sd_typeHex..trend                 0.287   NA           NA   trend   
## 4 cor_(Intercept).typeHex..trend   -0.989   NA           NA   trend   
## 5 sd_Observation.Residual           0.445   NA           NA   Residual
\end{verbatim}

\begin{verbatim}
##      
##        No Yes
##   No    0   0
##   Yes 673 431
\end{verbatim}

\includegraphics{paper_files/figure-latex/unnamed-chunk-11-1.pdf}

@@

\begin{verbatim}
## Linear mixed model fit by REML ['lmerMod']
## Formula: detect ~ type * trend + (1 | contributor)
##    Data: d
## REML criterion at convergence: 1350.719
## Random effects:
##  Groups      Name        Std.Dev.
##  contributor (Intercept) 0.1179  
##  Residual                0.4296  
## Number of obs: 1104, groups:  contributor, 92
## Fixed Effects:
##                (Intercept)                    typeHex.  
##                     0.5054                      0.1033  
##          trendthree cities             trendall cities  
##                    -0.4674                     -0.2772  
## typeHex.:trendthree cities    typeHex.:trendall cities  
##                     0.2500                      0.2391
\end{verbatim}

@@

\hypertarget{overall-performance}{%
\subsection{Overall Performance}\label{overall-performance}}

@@

\includegraphics{paper_files/figure-latex/demogs-1.pdf}

\begin{verbatim}
## # A tibble: 2 x 2
##   gender     n
##   <chr>  <int>
## 1 He        67
## 2 She       25
\end{verbatim}

\begin{verbatim}
## # A tibble: 3 x 2
##   education       n
##   <chr>       <int>
## 1 Bach           56
## 2 High School    23
## 3 Masters        13
\end{verbatim}
